Immutable Source-Mapped Abstract Syntax Tree:\ \\ A Design Pattern for Refactoring Engine APIs

### 2016-07-26 paper.tex
Cleaned up a few spots

\item The API should be usable when building refactorings for large code bases.
Automated refactorings are most useful for maintaining such codes.

@@

Mutable ASTs are conceptually simple and allow code for specific refactorings
to be very short. Performing refactorings on source code is conceptually
identical to changing the AST itself. This can greatly lower the bar for 
designing refactorings. However, mutable ASTs introduce a host of problems when
used in a large API:

@@

Guidelines for API design are discussed in detail in Jaroslav Tulach's
\textit{Practical API Design}~\cite{tulach08practical}.  A presentation by
Joshua Bloch~\cite{bloch} offers excellent advice as well.  Both authors
suggest minimizing the amount of mutability in an API.  Immutable objects tend
to have simpler interfaces, and they are thread-safe.  Both authors also
emphasize that APIs should be easy to learn (even for programmers who do not
read all of the API documentation) and difficult for clients to misuse. Mutable
ASTs have the advantage of ease-of-learning, but have have all the disadvantages
listed above. They can be valid solutions for simpler refactorings or projects
developed by a small team of people, but become problematic when used for 
nontrivial refactorings or for a public API. For a general, public refactoring
API, immutable trees are much more reliable.

@@

    \item \textit{Work with the ASTs in a functional style, where every edit
        returns a new AST.} So, a call to field.removeFromTree() would not
        change the existing AST, but rather it would return a new AST that has
        the given field omitted. When dealing with immutable data structures,
        functional programming may often be the natural solution. However,
        in the case of ASTs, a functional style risks reintroducing many of 
        the problems inherent to mutable trees. File location of nodes may
        be disrupted, pointers to nodes in the original tree may be rendered 
        irrelevant, semantic information may be invalidated, etc. Those are in
        addition to the difficulties of functional programming, such as memory
        consumption and copying, especially with very large ASTs. This is
        a valid refactoring API design pattern, but any implementation using 
        this variant must carefully consider these problems.

@@

Both Clang and the original refactoring engine in Xcode~3.0 modify source code
using offset-length information directly.  Notably, Clang offers a
\textit{RewriteRope} class---essentially, a custom string class that allows for
efficient mid-string insertions and
deletions\footnote{\url{http://clang.llvm.org/doxygen/classclang_1_1RewriteRope.html}}.
(It uses a B-tree structure inspired by, but not identical to, that described
in the original paper on Ropes~\cite{boehm95ropes}.)

>>>>

\item The API should be usable when building refactorings for large code bases.
Automated refactorings are most useful for maintaining such code.

@@

Mutable ASTs are simple to understand and allow code for specific refactorings
to be very short. Performing refactorings on source code is conceptually
identical to changing the AST itself. This can greatly lower the bar for 
designing refactorings. However, mutable ASTs introduce a host of problems when
used in a large API:

@@

Guidelines for API design are discussed in detail in Jaroslav Tulach's
\textit{Practical API Design}~\cite{tulach08practical}.  A presentation by
Joshua Bloch~\cite{bloch} offers excellent advice as well.  Both authors
suggest minimizing the amount of mutability in an API.  Immutable objects tend
to have simpler interfaces, and they are thread-safe.  Both authors also
emphasize that APIs should be easy to learn (even for programmers who do not
read all of the API documentation) and difficult for clients to misuse. Mutable
ASTs have the advantage of ease-of-learning, but have have all the disadvantages
listed above. They can be valid solutions for simpler refactorings or projects
developed by a small team of people, but become problematic when used for 
nontrivial refactorings or for a public API. For a general, public refactoring
API, immutable trees are much more robust.

@@

    \item \textit{Work with the ASTs in a functional style, where every edit
        returns a new AST.} So, a call to field.removeFromTree() would not
        change the existing AST, but rather it would return a new AST that has
        the given field omitted. When dealing with immutable data structures,
        functional programming may often be the natural solution. However,
        in the case of ASTs, a functional style risks reintroducing many of 
        the problems inherent to mutable trees. File location of nodes may
        be disrupted in the new tree, pointers to nodes in the original tree ma
        be rendered 
        irrelevant, semantic information may be invalidated, etc. Those are in
        addition to the difficulties of functional programming, such as memory
        consumption and copying, especially with very large ASTs. This is
        a valid refactoring API design pattern, but any implementation using 
        this variant must carefully consider these problems.

@@

Both Clang and the original refactoring engine in Xcode~3.0 modify source code
using offset-length information directly.  Notably, Clang offers a
\textit{RewriteRope} class---essentially, a custom string class that allows for
efficient mid-string insertions and
deletions\footnote{RewriteRope Class Reference: \url{http://clang.llvm.org/doxygen/classclang_1_1RewriteRope.html}}.
(It uses a B-tree structure inspired by, but not identical to, that described
in the original paper on Ropes~\cite{boehm95ropes}.)
